%% ── Abstract ──────────────────────────────────────────────────────────────
\begin{abstract}
This paper presents the design and implementation of the Smart Campus
Classroom Management System, an IoT and edge-AI platform developed at
SMU\,--\,Mediterranean Institute of Technology.  The system addresses
three core problems in traditional classroom operation: manual attendance
taking (up to 15~minutes per lecture), the absence of real-time
environmental monitoring, and the lack of early-warning mechanisms for
academically at-risk students.

The solution integrates an ESP32 microcontroller---equipped with
temperature, humidity, air-quality, and sound sensors---with a Raspberry
Pi~4B edge server that runs all services without cloud dependency.
Student attendance is automated via camera-based face recognition using
DeepFace/FaceNet 128-dimensional embeddings with cosine-distance
matching.  Sensor data is streamed over MQTT (Mosquitto broker) to a
FastAPI backend and forwarded in real time to a React professor dashboard
via WebSocket.  AC and lighting are controlled automatically based on
configurable thresholds, with manual professor override.  Attendance
records are synchronised to the Moodle LMS on session completion.

An AI analytics layer, built on a locally hosted Ollama phi3-mini LLM,
identifies at-risk students (below 70\% attendance) and generates
natural-language explanations; a deterministic forecasting pipeline
classifies per-course attendance trends and adds LLM-generated prose
interpretations.  The complete stack is containerised with Docker~Compose
and was delivered across 22 development phases within a single academic
semester.  Evaluation demonstrates full elimination of manual attendance
overhead, end-to-end sensor latency below 5~seconds, and AI pipeline
runtime reduction from approximately 14--16~minutes to 2--3~minutes
through batch-query optimisation.
\end{abstract}

%% ── Keywords ──────────────────────────────────────────────────────────────
\begin{IEEEkeywords}
IoT, smart classroom, face recognition, MQTT, FastAPI, edge computing,
LLM, at-risk detection, attendance automation, Raspberry Pi, ESP32,
Docker, WebSocket, Moodle
\end{IEEEkeywords}
